---
title: "EDA for Traffic Announcements"
author: "Group A.K.A.-R : Aybike Dilek / Serdar Kami Uckardes / Rabia Betul Parlar/ Alican Sevimay"
output: 
  ioslides_presentation:
    xaringan::moon_reader:
       lib_dir: libs
       css: [default, rladies-fonts, rladies]
       nature:
          highlightStyle: github
          highlightLines: true
          countIncrementalSlides: false
    fig_height: 24
    fig_width: 32
    widescreen: yes
     


---

```{css, echo = FALSE}
<style>
.column-left{
  float: left;
  width: 50%;
  text-align: left;
}
.column-center{
  display: inline-block;
  width: 50%;
  text-align: center;
}

.column-right{
  float: right;
  width: 50%;
  text-align: right;
}



h1.title{
  color:#dc791c;
}
h2{
  color:#084a4a;
}
h3{
  color:#084a4a;
}
h4{
  color:#084a4a;
}

```


```{r setup, include=FALSE}
library(showtext)
library(xaringanthemer)
library(tidyverse)
library(readxl)
library(lubridate)
library(data.table)
library(leaflet)
library(kableExtra)

knitr::opts_chunk$set(echo = TRUE)
```
## Table of Contents

- Preprocessing

- Identifying Most Problematic Announcement Type

- Relationship Between the 2019 Election and Running Projects

- Effect of the Pandemic on the Number of Accidents

- Distribution of the Announcement Types During Holidays

- Conclusions

- References


## Preprocessing

- This dataset contains set of traffic announcements entered into the system by the Transport Management Center operators from **2018** to **2020** which are stored in **2** different datasets.  
- There are **12** variables, with using Rbind **58,422** observations is obtained.
- We converted Turkish variables and record names into English words.
- We transformed **Announcement Types** and **Accident Types** variables into factorized variables.
- We filtered out some observations, where we did some calculations, in the 2020 dataset because of the change in date and time structure but kept them in time series operations where we just used count operation.



<font size="6">Identifying Most Problematic Announcement Type</font>
-----------------------------------------------------------------------
```{r,warning=FALSE,message=FALSE,echo = FALSE }
DS_1 <- read_excel("2018-2019-yl-ulam-yonetim-merkezi-tarafndan-sisteme-girilen-trafik-duyurular.xlsx")
DS_2 <- read_excel("2020-yl-ulam-yonetim-merkezi-tarafndan-sisteme-girilen-trafik-duyurular (1).xlsx")
DS=rbind(DS_1,DS_2)

colnames(DS)[1]='Announcement_ID'
colnames(DS)[2]='Announcement_Types'
colnames(DS)[3]='Announcement_Title'
colnames(DS)[4]='Announcement_Text'
colnames(DS)[5]='Starting_Time'
colnames(DS)[6]='Ending_Time'
colnames(DS)[7]='Intervention_Time'
colnames(DS)[8]='Accident_Situation'
colnames(DS)[9]='Effected_Lanes'
colnames(DS)[10]='Coordinates'
colnames(DS)[11]='Location'
colnames(DS)[12]='Accident_Types'

DS$Announcement_Types <- factor(DS$Announcement_Types,levels = c("Kaza Bildirimi", "Yoğun Trafik", 
                                                             "Araç Arızası", "Haber",
                                                             "Bakım-Onarım Çalışması",
                                                             "Çevre Düzenlemesi", "Araç Yangını",
                                                             "Yolun Trafiğe Kapanması",
                                                             "Şehir Hatları Sefer Bildirisi",
                                                             "İDO Sefer Bildirisi","Yol Yapım Çalışması",
                                                             "Alt Yapı Çalışması",
                                                             "Yolu Etkileyen Hava Koşulu",
                                                             "İmalat Çalışması","Diğer",
                                                             "Kontrollü Kapama (Protokol)","NULL"))
levels(DS$Announcement_Types) <-c("Accident Notification","Traffic Jam","Vehicle Breakdown","News","Maintenance",
                                  "Landscaping","Vehicle Fire","Road Closure","Marine Traffic Information-State",
                                  "Marine Traffic Information-Private","Construction Work-Road","Infrastructure Work",
                                  "Weather Condition","Construction Work-Structure","Others","Road Closure-Protocol","NULL")


DS$Accident_Types <- factor(DS$Accident_Types,levels = c("trafik kazası (hasarlı)", "trafik kazası (yaralanmalı)", 
                                                         "trafik kazası (zincirleme)", "trafik kazası (can kaybı)",
                                                         "Seçiniz")) 
levels(DS$Accident_Types) <- c("Vehicle Damage Only", "w/ Injuries","Pile-up Accident", "w/ Fatality","Others")

P1<- DS %>% mutate(Diff_Time_Hours = round(difftime(DS$Ending_Time,DS$Starting_Time)/3600,digits = 2))
P1 <- P1 %>% mutate(Problem_Index = P1$Diff_Time_Hours*P1$Effected_Lanes)
P1$Diff_Time_Hours = as.numeric(P1$Diff_Time_Hours)
P1$Problem_Index = as.numeric(P1$Problem_Index)

P1_1 <- P1 %>% filter(Problem_Index >= 0) %>% 
  filter(Problem_Index<= 3*sd(Problem_Index)+mean(Problem_Index) )%>%
  group_by(Announcement_Types) %>% summarise(Avg_Problem_Index = round(mean(Problem_Index),2),
  CountAnnouncement= n(), SumIndex=sum(Problem_Index))%>% filter(Avg_Problem_Index > 0) 
P1_1 <- data.frame(P1_1)
P1_1<- P1_1[order(P1_1$Avg_Problem_Index,decreasing = TRUE), ]
P1_1
```

- **Problem Index = Duration x Effected Lanes**

- <font size="4">On average, the most problematic announcement type was the **Infrastructure Work**.</font>

- <font size="4">In total, **maintenance** and **accident notification** made much more traffic since they occurred more frequently.</font>



<font size="6">Identifying Most Problematic Announcement Type</font> 
-----------------------------------------------------------------------
```{r,warning=FALSE,message=FALSE,echo = FALSE,fig.height = 5, fig.width = 8,fig.align= "center"}


ggplot(P1_1,aes(x="", y=Avg_Problem_Index, fill=Announcement_Types)) +geom_bar(stat='identity',width = 1)+coord_polar("y")+labs(title = "The Distribution of the Announcement Types Regarding to the Avg Problem Index")

```

<font size="6">Identifying Most Problematic Five Locations</font> 
-----------------------------------------------------------------------
```{r,warning=FALSE,message=FALSE,echo= FALSE}
P1_2 <- P1 %>% filter(Problem_Index >= 0) %>% select(Location,Problem_Index,Coordinates)
P1_2 <- data.frame(P1_2)
P1_2 <- P1_2[order(P1_2$Problem_Index,decreasing = TRUE), ]
P1_2 <- head(P1_2,5)
P1_2 <- data.frame(P1_2,do.call(rbind,strsplit(P1_2$Coordinates,",")))
P1_2$X2<- as.character(as.factor(P1_2$X2))
P1_2$X2 = as.numeric(P1_2$X2)
P1_2$X1<- as.character(as.factor(P1_2$X1))
P1_2$X1 = as.numeric(P1_2$X1)
P1_2  %>% select(Location,Problem_Index)
```

<font size="6">Identifying Most Problematic Five Locations on the Map</font> 
-----------------------------------------------------------------------
```{r,warning=FALSE,message=FALSE,echo= FALSE}

leaflet(P1_2) %>% 
  setView(lat = 33 ,lng = 40 , zoom = 7.1)  %>%  addTiles() %>% addCircles(data = P1_2, lat = ~ X1, lng = ~ X2, weight = 1,
                                                                             radius = ~sqrt(Problem_Index)*5,
                                                                             popup = ~as.character(Location), 
                                                                             color = "#d3501d", 
                                                                             fillOpacity = 20)




```


<font size="6">Relationship Between the 2019 Election and Running Projects</font>
-----------------------------------------------------------------------
```{r,warning=FALSE,message=FALSE,echo=F, results='hide',fig.height = 3.5, fig.width = 10,fig.align= "center"}
P2 = DS %>% filter(Announcement_Types %in% c('Landscaping','Maintenance','Infrastructure Work','Construction Work-Structure','Construction Work-Road')& Effected_Lanes >= 0)
P2_1 = P2 %>% group_by(Year_Quarter= lubridate::quarter(Starting_Time, with_year = TRUE)) %>% summarise(n = n())
P2_1$Year_Quarter = factor(P2_1$Year_Quarter)
P2_1=P2_1 %>% filter(Year_Quarter != "2020.4")

ggplot(P2_1, aes(x=Year_Quarter, y=n)) + geom_bar(stat="identity", position="dodge",fill="#d3501d", colour="black")+ theme_minimal()+ labs(title = "The Number of Running Projects on a Quarterly Basis",subtitle ="From 2018 to 2020")

```

- <font size="3">The first round of the election in March 2019</font>
- <font size="3">The second round of the election in June 2019</font>
- <font size="3">Filtered announcement types are; Landscaping, Maintenance, Infrastructure Work, Construction Work-Structure, Construction Work-Road.</font>


<font size="6">Effect of the Pandemic on the Number of Accidents</font>
-----------------------------------------------------------------------

```{r,warning=FALSE,message=FALSE,echo=F, results='hide',fig.height = 3.5, fig.width = 10,fig.align= "center"}

P3=setDT(DS)[, Year_Month := format(as.Date(`Starting_Time`), "%Y-%m") ]
P3= P3 %>% mutate(Year=year(Starting_Time),Day=day(Starting_Time))
P3_1 = P3 %>% filter(Announcement_Types == "Accident Notification" & Effected_Lanes >= 0 & Year_Month != "2020-11") %>% group_by(Year_Month,Year) %>% summarise(n = n())
P3_1$Year=as.factor(P3_1$Year)

ggplot(P3_1,aes(x=Year_Month,y = n)) + geom_bar(stat="identity",aes(fill=Year))+theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title = "The Number of Accidents on a Year-Month Basis",subtitle ="From 2018 to 2020")

```


- <font size="4">The first pandemic case occurred in March in Turkey.</font> 

- <font size="4">In April and May, there were a lot of lockdowns were applied.</font> 

<font size="6">Daily Number of the Accidents (2018-07)</font>
-----------------------------------------------------------------------
```{r,warning=FALSE,message=FALSE,echo=FALSE,fig.height = 3.5, fig.width = 10,fig.align= "center"}
P3_2 = P3 %>% filter(Announcement_Types == "Accident Notification" & Effected_Lanes >= 0 & Year_Month == "2018-07") %>% group_by(Day) %>% summarise(n = n())
P3_2$Day=as.factor(P3_2$Day)
ggplot(P3_2,aes(x=Day,y = n)) + geom_bar(stat="identity",fill="#FFC300", colour="black")+theme(axis.text.x = element_text(angle = 90, hjust = 1)) + geom_hline(yintercept=45, color="red")+
labs(title = "The Number of Accidents on a Daily Basis",subtitle = "July,2018")

```

- <font size="4">15’th of July is a national holiday in Turkey.</font>

<font size="6">Distribution of the Announcement Types During Holidays</font>
-----------------------------------------------------------------------
<div class="columns-2">

<div class="column-left">

<font size="4">**Holiday event calendar is introduced to the dataset.**</font>

<font size="4">During Holiday;</font>

- <font size="4">Accident and vehicle breakdown reasons were the majority of announcements.</font>

- <font size="4">The municipality may prefer to work on maintenance works.</font>

</div>


<div class="column-center">
```{r,warning=FALSE,message=FALSE,echo=FALSE,fig.height = 5, fig.width = 5 }

holidayCalendar <- read_excel("CalendarEvents.xlsx")
P4 <- DS %>%
mutate(weekStarting_Time = lubridate::week(DS$Starting_Time), yearDS =
lubridate::year(DS$Starting_Time))
P4_1 <- mutate(P4, holidayWeek =
(P4$weekStarting_Time==holidayCalendar$WeekHolidayStartDate &
P4$yearDS==holidayCalendar$Year |
P4$weekStarting_Time==holidayCalendar$WeekHolidayEndDate &
P4$yearDS==holidayCalendar$Year)) %>%
filter(Effected_Lanes >= 0 & holidayWeek==TRUE) %>%
group_by(Announcement_Types) %>% count(Announcement_Types)

ggplot(P4_1,aes(x="", y=n, fill=Announcement_Types)) +geom_bar(stat='identity',width =
1)+coord_polar("y")+scale_fill_hue(l=85)+labs(title = "The Distribution of Announcement Types During Holiday Weeks")
```

</div>

</div>

<font size="6">Conclusions</font>
-----------------------------------------------------------------------

- <font size="4">According to our problem index, infrastructure work & construction work-road were the worst traffic makers.</font>

- <font size="4">Maintenance & Accident Notification occurred more frequently, with this aspect,  these were the worst for traffic.</font>

- <font size="4">Considering that there will be an election in March 2019, the number of projects carried out compared to the last quarter of the previous year has increased.</font>

- <font size="4">With the new administration and the elimination of election uncertainty, the number of projects carried out by the municipality has started to increase regularly from the last quarter of 2019.</font>

- <font size="4">The number of accidents significantly decreased during the pandemic.</font>

- <font size="4">After rule bendings, the number of accidents increased to the standard levels.</font>

- <font size="4">During holiday times, accident notifications & vehicle breakdown notifications composed about 80% of total notifications.</font>

<font size="6">References</font>
-----------------------------------------------------------------------

- [Data Source - İBB Open Data Portal](https://data.ibb.gov.tr/dataset/ulasim-yonetim-merkezi-uym-tarafindan-sisteme-girilen-trafik-duyurulari)

- [External Data Source - Holiday Events](https://www.tatilsepeti.com/resmi-tatil-gunleri)

- [Mapping Link](https://medium.com/@joyplumeri/how-to-make-interactive-maps-in-r-shiny-brief-tutorial-c2e1ef0447da)


<font size="4">*You may click* [here](https://pjournal.github.io/mef04g-a-k-a-r/) *to reach other items of A.K.A - R Group’s Progress Journal.*</font>

```{r}



